---
title: "Ame Ind"
author: "LANG LI"
date: "July 28, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(DataComputing)
EMSdata = read.csv("EMSdata.csv")
AI = subset(EMSdata,EMSdata$E6_12_Race1 == "American Indian")
AIN = subset(AI,AI$Naloxone>0)
```

```{r,include=FALSE}
percentagePie = function(data,var,title){
  alcdrug = as.data.frame(table(data[,var]))
  lbls = alcdrug$Var1
  slices = alcdrug$Freq
  pct <- round(slices/sum(slices)*100)
  lbls <- paste(lbls, pct) # add percents to labels 
  lbls <- paste(lbls,"%",sep="") # ad % to labels 
  pie(slices,col=rainbow(length(lbls)),
  	main=title,labels = NA)
  legend(.9,.9,lbls,cex = 0.5,fill =rainbow(length(lbls)))
}
```

```{r,include=FALSE}
plotData = function(data){
  oddf = as.data.frame(table(data$E6_11_Gender1,data$agegroup))
  colnames(oddf) = c("Gender","Agegroup","Freq")
  oddf$Agegroup<-factor(oddf$Agegroup,levels = c("unknown","1-4","5-9","10-14","15-19","20-24","25-29","30-34","35-39","40-44","45-49","50-54","55-59","60-64","65-69","70-74","75-79","80-84","85+"))
  ggplot(oddf,aes(x=Agegroup,y=Freq ,fill=Gender))+geom_bar(stat='identity',position='dodge', width=.9) + theme(axis.text.x=element_text(angle=60,hjust=1))+labs( y = "Number of Cases")
}
```
## 1. Who (American Indian) 

Among the 2380 cases we have for American Indian patients. only 92 people received Naloxone

## 2. Where

```{r}
percentagePie(AI,36,"Pie Chart of the Urbanicity of American Indian patient")
```

66% of the patients were from Urban area while there are 24% from wilderness area. I think this has something to do with the accessibility of naloxone in wilderness area. 1435 counts (60.3%) were from Hennepin County, 312 counts were from Beltrami County (13.1%), 110 cases were from Cass county(4.6%) and Mille Lacs County has about 4%.

```{r}
percentagePie(AIN,36,"Pie Chart of the Urbanicity of American Indian patient")
```

## 3. Naloxone administration

1930 cases list drug overdose as primary symptom and 424 cases list drug overdose as secondary symptons. 26 cases did not involve any drug overdose symptons.

Looking at all chief complaints, there are 130 ETOH OR ALCOHOL overdose. 
256 case where total glasgow coma score <10, and 65 cases (25.4) where Naloxone is given.  18 of them have GCS <10. 59.3% of chief complaint is missing...

## 4. Demographics
```{r}
plotData(AI)
```
Most of the patients are male, bimodal distribution with peaks at 30-34 and  at 45-54. Female are relatively constant with a peak at 45-49. when we look at the naloxone administration data...
```{r}
plotData(AIN)
```

The distribution is stunning different. Administration is preferred on young people and much more female get administered...