---
title: "Data Analysis 3"
author: "Lang Li"
date: "June 28, 2017"
output: html_document
---


This file used to do some data cleaning:

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
Naloxone = read.csv("Naloxone.csv")
NalDO = read.csv("NalDO.csv")
NalNotDO = read.csv("NalNotDO.csv")
drugoverdose = read.csv("drugoverdose.csv")
overdoseCP = read.csv("overdoseCP.csv")
noODCP = read.csv("noODCP.csv")
```

## 1. Number of times we administer naloxone for DO patients:
```{r}
barplot(table(Naloxone$NaloxoneUsed)/sum(Naloxone$NaloxoneUsed)*100,xlab = "Naloxone dosages",ylab = "percentage",main = "Naloxone usage for all patients")
barplot(table(NalDO$NaloxoneUsed)/sum(NalDO$NaloxoneUsed)*100,xlab = "Naloxone dosages",ylab = "percentage",main = "Naloxone usage for Drug overdose patients")
barplot(table(NalNotDO$NaloxoneUsed)/sum(NalNotDO$NaloxoneUsed)*100,xlab = "Naloxone dosages",ylab = "percentage",main = "Naloxone usage for Non Drug overdose patients")
prim = table(NalNotDO$E9_15_PrimaryImp1)
barplot(prim,xlab = "Sympton numbers",ylab = "number of cases",main = "Primary impression of Non-drug overdose patient with Naloxone use")
```

From the first three plot, we can see that the dosages (number of time EMS personnel use Naloxone) have fairly similar distribution.

## 2. Primary Impression and Secondary Impression
So based on String matching, which is actually very exhausting.... we found that there are 130 cases there the chief complaint mentions "drug overdose". Now, we are investigating the code:
```{r}
barplot(table(overdoseCP$E9_15_PrimaryImp1),xlab = "Sympton numbers",ylab = "number of cases",main = "Primary impression of Patients with chief complaint of overdose")
```

So most likely, the code will be 1630 (altered level of Consciousness) and 1735 (Syncope/Fainting). 
```{r}
barplot(table(overdoseCP$E9_16_SecondaryImp1),xlab = "Sympton numbers",ylab = "number of cases",main = "Secondary impression of Patients with chief complaint of overdose")
```

So missing data for secondary impression still exists and the most secondary impressions are 1770 and 1870. So 1770 is "psychiatric disorder" while 1870 stands for "syncope / fainting".

We may also see the Naloxone usage for these categories of patients
```{r}
barplot(table(overdoseCP$NaloxoneUsed)/sum(overdoseCP$NaloxoneUsed)*100,xlab = "Dosages",ylab = "Number of Cases",main = "Naloxone Dosage for suspected overdose patients")
```


## 3. Signs and Symptoms
```{r}
table(drugoverdose$E9_14_SignsSymptoms1)
```

From the table, we observed that most of the patients included in this dataset are recorded as "None". Few others mentions "change in responsiveness", "Mental/Psych", "pain" and "Weakness". Missing data and Non sympton are majority....

## 4. Primary Organ affected
```{r}
table(drugoverdose$E9_12_PrimaryOrganSys1)
```

Here again, we have the most count as "Global pain or other illness" which fails to give useful information...